### 2018-08-31 reports/esi2.tex

\documentclass[11pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[super,sort&compress,comma]{natbib}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[detect-all]{siunitx}
\usepackage{xr}
\externaldocument{apssamp}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\Roman{table}}
\author{Andrew R. McCluskey,\textit{$^{ab}$}$^{\ddag}$ Adrian Sanchez-Fernandez,\textit{$^{ac}$}$^{\ddag\P}$ \\
Karen J. Edler,\textit{$^{a}$}$^{\ast}$ Stephen C. Parker,\textit{$^{a}$} Andrew J. Jackson,\textit{$^{cd}$} \\
Richard A. Campbell,\textit{$^{ef}$} and Thomas Arnold\textit{$^{abcg}$}$^{\ast}$ \\
 \\
\textit{$^{a}$~Department of Chemistry, University of Bath, Claverton} \\
\textit{Down, Bath, BA2 7AY, UK.} \\
\textit{$^{b}$~Diamond Light Source, Harwell Campus, Didcot, OX11 0DE,} \\
\textit{UK.} \\
\textit{$^{c}$~European Spallation Source, SE-211 00 Lund, Sweden.} \\
\textit{$^{d}$~Department of Physical Chemistry, Lund University,} \\
\textit{SE-211 00 Lund, Sweden.} \\
\textit{$^{e}$~Institut Laue-Langevin, 71 avenue des Martyrs, 38000,} \\
\textit{Grenoble, France.} \\
\textit{$^{f}$~Division of Pharmacy and Optometry, University of} \\
\textit{Manchester, Manchester, M13 9PT, UK.} \\
\textit{$^{g}$~ISIS Neutron and Muon Source, Science and Technology} \\
\textit{Facilities Council, Rutherford Appleton Laboratory, Harwell} \\
\textit{Oxford, Didcot OX11 0QX, UK.} \\
\textit{$^{\P}$~Present address: Department of Food Technology, Lund} \\
\textit{University, SE-211 00 Lund, Sweden}}
\title{ESI for ``Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers''}
\date{}
\begin{document}

\maketitle

\section{Materials}
Choline chloride (99 \%, Sigma-Aldrich) and glycerol (99 \%, Sigma-Aldrich) d$_9$-choline chloride (99 \%, 98 \% D, CK Isotopes) and d$_8$-glycerol (99 \%, 98 \% D, CK Isotopes)  were purchased and used without further purification. The DES was prepared by mixing the precursors at the appropriate ratio, and heating at \SI{80}{\celsius} until a homogeneous, transparent liquid formed.\cite{Smith2014} The solvent was equilibrated overnight at \SI{40}{\celsius} and subsequently stored under a dry atmosphere. Due to the limited availability of the deuterated precursors, a fully protonated subphase (hDES) and a partially deuterated subphase (hdDES) were prepared and used during the neutron reflectometry (NR) experiment. The partially deuterated subphase was prepared using the following mixtures of precursors: \SI{1}{\mole} of 0.38 mole-fraction of h-choline chloride/0.62 mole-fraction of d-choline chloride; and \SI{2}{\mole} of 0.56 mole-fraction of h-glycerol/0.44 mole-fraction of d-glycerol. The solvent was subsequently prepared following the procedure discussed above.

The water content of the DES was determined before and after each experiment by Karl-Fischer titration (Mettler Toledo DL32 Karl-Fischer Coulometer, Aqualine Electrolyte A, Aqualine Catholyte CG A) in order to ensure water presence was kept to a minimum. Those measurements showed that the water content of the solvent was kept below \SI{0.3}{wt\percent} during all the experimental procedures presented here, which we assume to be negligible and have little impact on the characteristics of the DES.\cite{Hammond2016,Hammond2017}

DPPC (\SI{>99}{\percent}, C$_{16}$ tails), DMPC (\SI{>99}{\percent}, C$_{14}$ tails), and DMPG (\SI{>99}{\percent}, C$_{14}$ tails) were supplied by Avanti Polar Lipids and DLPC (\SI{>99}{\percent}, C$_{12}$ tails) was supplied by Sigma-Aldrich and all were used as received. Deuterated versions of DPPC (d$_{62}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) and DMPC (d$_{54}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification. These phospholipids were dissolved in chloroform (\SI{0.5}{\milli\gram\per\milli\liter}) at room temperature. PC indicates the molcule contains a phosphocholine head component, where PG contains a phosphoditylglycerol head component, these are shown in Figure \ref{fig:heads}.

In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a manageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controlable barriers. So, although the surface coverage was nominally the same as used in the X-ray studies, the lack of precise control over the surface pressure meant that it was not appropriate to co-refine XRR and NR contrasts together.

\section{Methods}
XRR measurements were taken on I07 at Diamond Light Source, at \SI{12.5}{\kilo\eV} photon energy using the double-crystal-deflector.\cite{Arnold2012} The reflected intensity was measured in a momentum transfer range from \SIrange{0.018}{0.7}{\AA^{-1}}. The data were normalised with respect to the incident beam and the background was measured from off-specular reflection and subsequently subtracted. Samples were equilibrated for at least one hour and preserved under an argon atmosphere to minimise the adsorption of water by the subphase. XRR data were collected for each of the lipids, DLPC, DMPC, DPPC and DMPG at four surface pressures (DLPC: 20, 25, 30, and \SI{35}{\milli\newton\meter^{-1}}, DMPC: 20, 25, 30, and \SI{40}{\milli\newton\meter^{-1}}, DPPC: 15, 20, 25, and \SI{30}{\milli\newton\meter^{-1}}, DMPG: 15, 20, 25, and \SI{30}{\milli\newton\meter^{-1}}), as measured with an aluminium Wilhelmy plate; all measurements were conducted at \SI{22}{\celsius}. The aluminium Wilhelmy plate was used over a traditional paper plate due to the low wetability of paper by the DES.

The NR experiments were performed on FIGARO at the Institut Laue-Langevin using the time-of-flight method.\cite{Campbell2011} Data at two incident angles of \SI{0.62}{\degree} and \SI{3.8}{\degree} were measured to provide a momentum transfer range from \SIrange{0.005}{0.18}{\AA^{-1}}. Two surface pressures for each system and contrast was measured (DMPC: 20 and \SI{25}{\milli\newton\meter^{-1}}, DPPC: 15 and \SI{20}{\milli\newton\meter^{-1}}). Similar to the X-ray procedure, samples were given enough time to equilibrate (at least two hours), kept under an inert atmosphere, and all measurements were conducted at \SI{22}{\celsius}.

\section{Literature values for head and tail volumes}
\begin{table}
	\caption{\ Lipid component volumes extracted from different literature sources. $V_l$ corresponds to the total lipid volume, MD to molecular dynamics simulation, WAXS to wide-angle X-ray scattering, NB to neutral buoyancy and DVTD to differential vibrating tube densimetry. $^a$ The values for the head component in Kucerka \emph{et al.},\cite{Kucerka2004} were taken from Balgav\'{y} \emph{et al}.\cite{Balgavy2001}}
  \centering
	\label{tab:water}
	\begin{tabular}{llllllllll}
		\hline
    Lipid & DPPC & & & DMPC & & DLPC & & DMPG & POPG \\
    \hline
    Reference & [\cite{Armen1998}] & [\cite{Sun1994}] & [\cite{Kucerka2004,Balgavy2001}]$^a$ & [\cite{Armen1998}] & [\cite{Kucerka2004,Balgavy2001}]$^a$ & [\cite{Armen1998}] & [\cite{Kucerka2004,Balgavy2001}]$^a$ & [\cite{Pan2012}] & [\cite{Kucerka2012}] \\
    \hline
    $V_l$/\si{\angstrom^3} & $1287.3\pm25.5$ & $1148\pm2$ & $1264.2\pm32.1$ & $1172.5\pm25.1$ & $1155.4\pm30.0$ & $1057.7\pm24.7$ & $1046.6\pm28.0$ & $1011.4$ & $1203$ \\
    $V_t$/\si{\angstrom^3} & $966.4\pm5.4$ & $829\pm4$ & $924.7\pm17.6$ & $851.5\pm5.0$ & $815.9\pm15.5$ & $736.8\pm4.6$ & $707.1\pm13.5$ & $720.4$ & $914$ \\
    $V_h$/\si{\angstrom^3} & $320.9\pm20.1$ & $319\pm6$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $291.0$ & $289$ \\
    Method & MD & WAXS & NB & MD & NB & MD & NB & DVTD & MD \\
    T/\si{\celsius} & 50 & 24 & 30 & 50 & 30 & 50 & 30 & 20 & 25 \\
	\end{tabular}
\end{table}

\section{Chemically-consistant model}
Our chemically-consistent model is made up of two layers to define the lipid monolayer; the head layer at the interface with the solvent and tail layer at the interface with the air. The head components have a calculated scattering length, $b_h$, (found as a summation of the X-ray or neutron atomic scattering lengths), and a component volume, $V_h$. These head components make up a layer with a given thickness, $d_h$, and roughness, $\sigma_h$, within which some volume fraction of solvent can intercalate, $\phi_h$. The tail components also have a calculated scattering length, $b_t$, roughness, $\sigma_t$, and a component volume, $V_t$. We have defined the thickness of the tail layer, $d_t$, in terms of the maximum (all-trans) length of the carbon tail, $t_t$, and the angle that the chain is tilted by with respect to the interface normal, $\theta_t$,
%
\begin{equation}
\label{equ:tl}
d_t = t_t \cos{\theta_t}.
\end{equation}
%
This was used to impose a chemically-sensible maximum on the thickness of the lipid tail layer, e.g. it cannot be thicker than the maximum extended length of the lipid tail. The scattering length density (SLD) of the tail and head layers used in the Abel\`{e}s model can therefore be found as follows,
%
\begin{equation}
\text{SLD}_i = \frac{b_i}{V_i}(1 - \phi_i) + \text{SLD}_{s}(\phi_i),
\end{equation}
%
where, $\text{SLD}_{s}$ is the scattering length density of the subphase (DES), and $i$ indicates either the tail or head layer; it is assumed that the tail layer contains neither solvent nor air, e.g. $\phi_t = 0$. To ensure that the number density of head components and pairs of tail components is the same, the following constraint was included in the model,\cite{Braun2017}
%
\begin{equation}
\label{equ:phih}
\phi_h =  1 - \bigg(\frac{d_tV_h}{V_td_h}\bigg).
\end{equation}
%
Based on the work of Campbell \emph{et al.},\cite{Campbell2018} a single value for the interfacial roughness was fitted for all of the interfaces, including the subphase (i.e. $\sigma_h$ = $\sigma_t$ = $\sigma_s$), as there is only a single lipid molecule type in each monolayer. Therefore, any capillary wave roughness at the air-DES interface is carried equally through the layers.

%
\begin{table}[h]
	\small
	\caption{\ The invariant parameters within the chemically-consistent model.
	$^a$Values obtained from the Tanford formula.\cite{Tanford1980} $^b$Values obtained from Sanchez-Fernandez \emph{et al.}\cite{Sanchez-Fernandez2016}}
	\label{tab:invariant}
  \centering
	\begin{tabular}{ccccc}
		\hline
		Component & $b_t$/fm & $b_h$/fm & $t_t$/\AA & $\text{SLD}$/$\times10^{-6}$\AA$^{-2}$ \\
		\hline
		X-ray & & & & \\
		DLPC & 5073 & 4674 & 15.5$^a$ & -- \\
		DMPC & 5985 & 4674 & 18.0$^a$ & -- \\
		DPPC & 6897 & 4674 & 20.5$^a$ & -- \\
		DMPG & 5985 & 4731 & 18.0$^a$ & --\\
		Air & -- & -- & -- & 0\\
		DES & -- & -- & -- & 10.8$^b$ \\
		\hline
		Neutron & & & & \\
		d$_{54}$-DMPC & 5329.8 & 602.7 & 18.0$^a$ & -- \\
		d$_{62}$-DPPC & 6129.2 & 602.7 & 20.5$^a$ & -- \\
		h-DES & -- & -- & -- & 0.43$^b$  \\
		hd-DES & -- & -- & -- & 3.15$^b$ \\
		\hline
	\end{tabular}
\end{table}
%

\section{Grazing incident X-ray diffraction (GIXD)}
\label{sec:gixd}
GIXD was measured for DPPC and DMPC at \SI{30}{\milli\newton\per\meter} at \SI{22}{\celsius} (and \SI{7}{\celsius} for DMPC), and are shown in Figures \ref{fig:dppcgixd}-\ref{fig:dmpcgixd7}. All figures contain an artifact from the X-ray beam interacting with the Teflon of the Langmuir trough. However, in figures \ref{fig:dppcgixd} and \ref{fig:dmpcgixd7} it is possible to identify a (2, 0) diffraction peak that indicates the presence of a similar structure to that found under the same conditions in water.\cite{Watkins2009} Based on this information we believe that, similar to water, the phase of the lipid tails is likely to be in the liquid phase at all surface pressures measured.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/206041}
	\caption{The GIXD pattern for DPPC at \SI{30}{\milli\newton\per\meter} at \SI{22}{\celsius}, the axes at in units of \si{\per\angstrom}. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppcgixd}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/206151}
	\caption{The GIXD pattern for DMPC at \SI{30}{\milli\newton\per\meter} at \SI{22}{\celsius}, the axes at in units of \si{\per\angstrom}. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpcgixd}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/206161}
	\caption{The GIXD pattern for DMPC at \SI{30}{\milli\newton\per\meter} at \SI{7}{\celsius}, the axes at in units of \si{\per\angstrom}. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpcgixd7}
\end{figure}

\section{Probability distribution functions}

The two-dimensional probability distribution functions (PDFs) for all parameters and all lipids from the X-ray reflectometry models are given in Figures \ref{fig:dlpc2}-\ref{fig:dmpg5}. The two-dimensional probability distribution functions (PDFs) for all parameters and all lipids from the neutron reflectometry models are given in Figures \ref{fig:dmpcn1}-\ref{fig:dppcn2}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dlpc2_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DLPC X-ray reflectometry data at 20 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dlpc2}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dlpc3_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DLPC X-ray reflectometry data at 25 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dlpc3}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dlpc4_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DLPC X-ray reflectometry data at 30 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dlpc4}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dlpc5_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DLPC X-ray reflectometry data at 35 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dlpc5}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpc2_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPC X-ray reflectometry data at 20 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpc2}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpc3_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPC X-ray reflectometry data at 25 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpc3}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpc4_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPC X-ray reflectometry data at 30 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpc4}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpc5_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPC X-ray reflectometry data at 40 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpc5}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dppc2_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DPPC X-ray reflectometry data at 15 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppc2}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dppc3_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DPPC X-ray reflectometry data at 20 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppc3}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dppc4_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DPPC X-ray reflectometry data at 25 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppc4}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dppc5_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DPPC X-ray reflectometry data at 30 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppc5}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpg2_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPG X-ray reflectometry data at 15 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpg2}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpg3_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPG X-ray reflectometry data at 20 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpg3}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpg4_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPG X-ray reflectometry data at 25 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpg4}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpg5_all_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of DMPG X-ray reflectometry data at 30 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpg5}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpc20_neutron_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of two contrast DMPC neutron reflectometry data at 20 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpcn1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dmpc25_neutron_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of two contrast DMPC neutron reflectometry data at 25 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dmpcn2}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dppc15_neutron_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of two contrast DPPC neutron reflectometry data at 15 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppcn1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.50\textwidth]{figures/dppc20_neutron_corner}
	\caption{The multi-parameter PDFs for the chemically-relevant model of two contrast DMPC neutron reflectometry data at 20 mNm$^{-1}$. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:dppcn2}
\end{figure}

%%%REFERENCES%%%
\bibliography{rsc} %You need to replace "rsc" on this line with the name of your .bib file
\bibliographystyle{rsc} %the RSC's .bst file
\end{document}